LME-MPM applied to quasi-brittle fracture.

### 2020-05-21 2020_EFM_MPM_Eigensoftening.tex
Cambios 20-5-2020 (II) : Borrador terminado

\journal{Engineering Fracture Mechanics}

@@

The aim of this section is to provide an overview of the special
techniques employed to face the fracture problem under the \acrshort{mpm}
framework. In consequence it is structured as follows: first in
\ref{sec:2.1} the \acrfull{npc} algorithm for the \acrshort{mpm}
will be exposed, next the \acrshort{lme} approximants are
introduced in \ref{sec:2.2} as an accurate alternative technique to
interpolate data between particles and nodes, and finally fracture
algorithms based in the eigendeformation concept are presented in \ref{sec:2.3}.

\subsection{The \acrshort{mpm} time integration : A \acrlong{npc}  scheme}
\label{sec:2.1}
The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian
method where moving material points, henceforth particles, carries on
all the physical information of the local state ($\sigma_p,
\varepsilon_p$) and a set of fixed background nodes is employed to
compute the balance of momentum equation. Since the \acrshort{mpm}
possesses the advantages of both Lagrangian and Eulerian descriptions,
no element distortion takes place in the \acrshort{mpm}. Therefore it
is an appropriate and efficient method to solve problems with moving
discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation in a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research
a explicit predictor-corrector time integration scheme is adopted. The
purposes of this choice is motivated due its proved robustness and
stability for dynamic computations. In the first stage, the nodal
velocity predictor is computed following \eqref{eq:Predictor-velocity}, 
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{\tens{m}_I^{k+1}}\ .
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimise the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, a
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed thorough the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage,
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally updated the particles are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} =
          \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}
The complete pseudo-algorithm it is summarised in \ref{sec:expl-pred-corr}.

@@

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in both loading and fracture stages
in the reaction-displacements curve when linear interpolation technique is employed. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear. Besides that, linear interpolation
produces stiffer results than \acrshort{lme}. This can be attributed
to imprecision in the stress field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
On the other hand, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case. Although this behavior does not affect significantly to the peak
stress value in a hookean material \cite{Zhang_EE_2020}, if
hyperelastic materials are consider for future simulations,
inaccuracies in the stress field could affect dramatically to the
final result. On the other hand \acrshort{lme} produces soft stress
field evolution even after breaking.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated, Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver.

@@

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented. Several levels of discretization are employed to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, $\gamma$ was
fixed to 6.\\

First the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
supports. Therefore, this results can be improved in future research
by a suitable contact algorithm.\\

And finally horizontal stress distributions and crack propagation are
extracted, see Figure \ref{fig:Stress-vs-damage-impact-test}. Compared
with results obtained by \cite{Navas_2017_ES}, stress evolution

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:4}

@@

\section{Explicit Newmark Predictor-Corrector algorithm}
\label{sec:expl-pred-corr}
\begin{algorithm}[H]
  \DontPrintSemicolon
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% º
    \textbf{Update mass matrix :} $ \tens{m}_{I} = N_{Ip}^{k}\ m_p$ \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Predictor :}
    \begin{equation*}
      \vec{v}_I^{pred} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Impose essential boundary conditions :} At the fixed
    boundary, set $\vec{v}_{I}^{pred} = 0$.\; 
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Deformation tensor increment calculation :}
    \begin{equation*}
      \Delta \tens{\varepsilon}_{p}^{k+1} = \Delta t\
        \dot{\tens{\varepsilon}_{p}}^{k+1} = \Delta t\ \left[ \vec{v}_{I}^{pred} \otimes
        \Grad{N_{Ip}^{k+1}} \right]^s
    \end{equation*} \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Update the density field :} $\rho_p^{k+1} =
    \frac{\rho_p^k}{1 + \mathit{tra}\left[\Delta\tens{\varepsilon}_{p}^{k+1}\right]}.$\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Compute stress field and update damage parameter}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Balance of forces calculation :} Calculate the total grid
    nodal force $\vec{f}_{I}^{k+1} = \vec{f}_{I}^{int,k+1} + \vec{f}_{I}^{ext,k+1}$.\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Corrector :}
    \begin{equation*}
      \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}  
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Update particles lagrangian quantities :}
    \begin{align*}
      &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}
    \end{align*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Reset nodal values}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \label{alg-epc}
    \caption{Explicit Newmark Predictor-Corrector scheme}
\end{algorithm}

>>>>

\newcommand{\PNA}[1]{
  \textcolor{red}{{#1}}
}

\newcommand{\MMP}[1]{
  \textcolor{blue}{{#1}}
}

\newcommand{\DM}[1]{
  \textcolor{gree}{{#1}}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \journal{Engineering Fracture Mechanics}
\journal{Computers and Structures}

@@

The aim of this section is to describe and introduce some special
techniques required to face the fracture problem under the \acrshort{mpm}
framework. In consequence it is structured as follows: first in
\ref{sec:2.1} the \acrfull{npc} algorithm for the \acrshort{mpm}
will be exposed, next the \acrshort{lme} approximants are
introduced in \ref{sec:2.2} as an accurate alternative technique to
interpolate data between particles and nodes, and finally fracture
algorithms based in the eigendeformation concept are presented in \ref{sec:2.3}.

\subsection{The \acrshort{mpm} time integration : A \acrlong{npc}  scheme}
\label{sec:2.1}
The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian
method where moving material points, henceforth particles, carries on
all the physical information of the local state ($\sigma_p,
\varepsilon_p$) and a set of fixed background nodes are introduced to compute the balance of momentum equation. Since the \acrshort{mpm} 
possesses the advantages of both Lagrangian and Eulerian descriptions,
no element distortion takes place in the \acrshort{mpm}. Therefore it
is an appropriate and efficient method to solve problems with moving
discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation in a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research
a explicit predictor-corrector time integration scheme is adopted. The
purposes of this choice is motivated due its proved robustness and
stability for dynamic computations. In the first stage, the nodal
velocity predictor is computed following \eqref{eq:Predictor-velocity}, 
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{\tens{m}_I^{k+1}}\ .
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimise the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, a
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed thorough the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage,
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally updated the particles are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} =
          \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}
The complete pseudo-algorithm it is summarised in \ref{sec:expl-pred-corr}.

@@

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in both loading and fracture stages
in the reaction-displacements curve when linear interpolation
technique is used. In contrast, \acrshort{lme} simulation does not
exhibit these spurious oscillations and remains linear. Besides that,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
On the other hand, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case. Although this behavior does not affect significantly to the peak
stress value in a hookean material \cite{Zhang_EE_2020}, if
hyperelastic materials are consider for future simulations,
inaccuracies in the stress field could affect dramatically to the
final result. On the other hand \acrshort{lme} produces soft stress
field evolution even after breaking.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated, Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver.

@@

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented. Several levels of discretization were required to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, $\gamma$ was
fixed to 6.\\

First the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. Therefore, this results can be improved in future research
by a suitable contact algorithm.\\

And finally horizontal stress distributions and crack propagation are
extracted, see Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. In one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. And in the other
hand there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:4}
Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.}
\cite{Zhang_EE_2020}, we have proposed two addition improvements which
solves the shortcomings presented in the original concept. First by
bring to the \acrshort{mpm} framework the concept of eigensoftening,
which allow to consider gradual failure process of each material
point. This concept can be compared with cohesive law in the context
of cohesive fracture. Furthermore, more elaborated damage curves can
be designed to describe the failure process in sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such carbon fiber. This leads a challenging
research topic. Secondly, \acrshort{lme} is employed to remove from
the solution those wiggles due to inaccuracies in the stress
integration. This interpolation technique could help to improve the
results in fracture propagation by adapting it thorough the
deformation gradient. \textit{A priori} this will enhance the already
impressive localisation properties of eigensoftening and eigenerosion
algorithms.

@@

\section{Explicit Newmark Predictor-Corrector algorithm}
\label{sec:expl-pred-corr}
\begin{algorithm}[H]
  \DontPrintSemicolon
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Update mass matrix :} $ \tens{m}_{I} = N_{Ip}^{k}\ m_p$ \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Predictor :}
    \begin{equation*}
      \vec{v}_I^{pred} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Impose essential boundary conditions :} At the fixed
    boundary, set $\vec{v}_{I}^{pred} = 0$.\; 
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Deformation tensor increment calculation :}
    \begin{equation*}
      \Delta \tens{\varepsilon}_{p}^{k+1} = \Delta t\
        \dot{\tens{\varepsilon}_{p}}^{k+1} = \Delta t\ \left[ \vec{v}_{I}^{pred} \otimes
        \Grad{N_{Ip}^{k+1}} \right]^s
    \end{equation*} \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Update the density field :} $\rho_p^{k+1} =
    \frac{\rho_p^k}{1 + \mathit{tra}\left[\Delta\tens{\varepsilon}_{p}^{k+1}\right]}.$\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Compute stress field and update damage parameter}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Balance of forces calculation :} Calculate the total grid
    nodal force $\vec{f}_{I}^{k+1} = \vec{f}_{I}^{int,k+1} + \vec{f}_{I}^{ext,k+1}$.\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Corrector :}
    \begin{equation*}
      \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}  
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Update particles lagrangian quantities :}
    \begin{align*}
      &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}
    \end{align*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Reset nodal values}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \label{alg-epc}
    \caption{Explicit Newmark Predictor-Corrector scheme}
\end{algorithm}
